---
title: "Lab 4"
author: "Christina Lin"
date: "9/17/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1: Read the data 
```{r}
library(data.table)
library(tidyverse)

if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")

```

# Step 2: Prepare the Data

Removing temperatures less than -17 
```{r}
met <- met[temp >= -17]
```

Make sure there are no missing data in the key variables
```{r}
met[,range(temp)]
met[,range(rh, na.rm = TRUE)]
met[,range(wind.sp, na.rm = TRUE)]
met[,range(vis.dist, na.rm = TRUE)]
met[,range(dew.point, na.rm = TRUE)]
met[,range(lat, na.rm = TRUE)]
met[,range(lon, na.rm = TRUE)]
met[,range(elev, na.rm = TRUE)]

#Changing 9999 in elev to NA
met[elev == 9999.0, elev := NA]
```

Generate a date variable using the functions as.Date() 
```{r}
met[, ymd := as.Date(paste(year, month, day, sep = "-"))]
```


Using the data.table::week function, keep the observations of the first week of the month.
```{r}
met[, table(week(ymd))]

met <- met[week(ymd) == 31]
```

Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.

```{r}
met_avg <- met[, .(
  temp = mean (temp, na.rm = TRUE),
  rh = mean (rh, na.rm = TRUE),
  wind.sp = mean (wind.sp, na.rm = TRUE),
  vis.dist = mean (vis.dist, na.rm = TRUE),
  dew.point = mean (dew.point, na.rm = TRUE),
  lat = mean (lat, na.rm = TRUE),
  lon = mean (lon, na.rm = TRUE),
  elev = mean (elev, na.rm = TRUE), USAFID
), by = "USAFID"]
```


Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees

```{r}
met_avg[lat >= 39.71 & lon <= -98.00, region := "Northwest"]
met_avg[lat < 39.71 & lon <= -98.00, region := "Southwest"]
met_avg[lat >= 39.71 & lon > -98.00, region := "Northeast"]
met_avg[lat < 39.71 & lon > -98.00, region := "Southeast"]

#alternative
#fifelse(lat >=39.71 & lon <=-98, "Northwest",
#        fifelse(lat < 39.71 & lon <=-98, "Southwest",
#                fifelse(lat >= 39.71 & lon > -98, "Northeast",
#                       fifelse(lat < 39.71 & lon >-98, "Southeast", NA_character_))))

met_avg[, table(region, useNA="always")]

```


Create a categorical variable for elevation as in the lecture slides

```{r}
met_avg[, elev_cat := fifelse(elev >252, "high", "low")]

```

