---
title: "Lab_6"
author: "Christina Lin"
date: "10/1/2021"
output: 
  github_document:
  html_document:
    html_preview:false
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```


```{r}
if (!file.exists("mtsamples.csv"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv",
    destfile = "mtsamples.csv",
    method   = "libcurl",
    timeout  = 60
    )
mtsamples <- read.csv("mtsamples.csv")

library(dplyr)
library(ggplot2)
library(tidyverse)
library(tidytext)
```



```{r}
mtsamples <- as_tibble(mtsamples)
```

# Question 1: What specialties do we have?
```{r}
specialties <- mtsamples %>%
  count(medical_specialty)

specialties %>%
  arrange(desc(n)) %>%
  top_n(15) %>%
  knitr::kable()

```

There are `r nrow(specialties)` specialties. Looking at the distribution:

```{r}
ggplot(mtsamples,aes(x = medical_specialty)) +
  geom_histogram(stat = "count") +
  coord_flip()
```

The specialties are not evenly distributed.



# Question 2: Most frequent words in transcription column

```{r, cache = TRUE}
mtsamples %>%
  unnest_tokens(output = token,input = transcription) %>%
  count(token) %>%
  top_n(20,n) %>%
  ggplot(aes(n, fct_reorder(token,n))) + 
  geom_col()
```

The word "patient" seems to be important, but we observe a lot of stop words. 


# Question 3: Removing the stop words and numbers

```{r}
mtsamples %>%
  unnest_tokens(output = token,input = transcription) %>%
  count(token, sort=TRUE) %>%
  anti_join(stop_words, by = c("token" = "word")) %>%
  filter(!grepl(pattern="^[0-9]+$", x= token)) %>%
  top_n(20,n) %>%
  ggplot(aes(n, fct_reorder(token,n))) + 
  geom_col()

#using anti_join() after the count is computationally faster
#the filter function is filtering expressions that start and end with a number using regular expression
```


"Patient" is the most common word, followed by "left" and "history". This could suggest that there is something wrong with the patient's left side of the body. Numbers were part of the highest frequency words, so they were removed. 



# Question 4: Tokenizing with ngrams

First, tokenizing by bi-grams.
```{r}
mtsamples %>%
  unnest_ngrams(output = bigram,input = transcription, n = 2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(20)%>%
  ggplot(aes(n, fct_reorder(bigram,n))) + 
  geom_col()
```

Using bigrams is not very informative. Thus, trying trigrams:


```{r}
mtsamples %>%
  unnest_ngrams(output = trigram,input = transcription, n = 3) %>%
  count(trigram, sort = TRUE) %>%
  top_n(20) %>%
  ggplot(aes(n, fct_reorder(trigram,n))) + 
  geom_col()
```


Using trigrams, it appears that there are a lot of phrases that associate with surgery and operations. 



# Question 5: Words that appear before and after "left"

```{r, cache = TRUE}
bigrams <- mtsamples %>%
  unnest_ngrams(ngram,transcription,n=2) %>%
  separate(ngram, into = c("word1","word2", sep = " "))

#Words appearing after "left"
bigrams %>%
  select(word1, word2) %>%
  filter(word1 == "left") %>%
  count(word2, sort = TRUE)%>%
  top_n(20) %>%
  knitr::kable(caption = "Words AFTER 'left'")

#Words appearing before "left"
bigrams %>%
  select(word1, word2) %>%
  filter(word2 == "left") %>%
  count(word1, sort = TRUE) %>%
  top_n(20) %>%
  knitr::kable()

```

Since we are looking at single words again, it is good idea to treat these as single tokens. Therefore, remove stop words and numbers from word1. Words appearing after 'left' do not seem to have this problem. 


```{r}
bigrams %>%
  filter(word2 == "left") %>%
  filter(!(word1 %in% stop_words$word) & (!grepl(pattern="^[0-9]+$", word1))) %>%
  count(word1, sort= TRUE) %>%
  top_n(20) %>%
  knitr::kable(caption = "Words BEFORE 'left'")
```



# Question 6: Which words used the most in each specialty?

```{r}
mtsamples %>%
  group_by(medical_specialty) %>%
  unnest_tokens(output = token,input = transcription) %>%
  filter(!(token %in% stop_words$word) & (!grepl(pattern="^[0-9]+$", token))) %>%
  count(token, sort = TRUE) %>%
  top_n(5) %>%
  arrange(medical_specialty, desc(n)) %>%
  knitr::kable()
```

