---
title: "Lab-3"
author: "Christina Lin"
date: "9/10/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Read the data

```{r}

#to test if a package exists and prevent code from failing, you can use the require function

if (!require(data.table)) {
  install.packages("data.table") }

library(data.table)

#only download the file if it doesn't exist already (be aware of where the R markdown file and data file are located)
if (!file.exists("met_all.gz")){
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)}

met <- data.table::fread("met_all.gz")
```

## Step 2: Checking Dimensions

```{r}
dim(met)
head(met)
tail(met)
```
## Step 3: Looking at the variables

```{r}
str(met)
```

## Step 4: Looking at key variables

```{r}
table(met$year)

table(met$month)

table(met$day)

table(met$hour)

```

Looking at continuous variables 

```{r}
summary(met$temp)

summary(met$elev)

summary(met$wind.sp)
```
Updating NAs in elevation

```{r}
met$elev[met$elev == 9999.0] <- NA
#met[elev == 9999.0,elev := NA] this is filtering using data table

summary(met$elev)
```
The highest elevation is `r max(met$elev, na.rm = TRUE)` meters. The lowest elevation is `r min(met$elev, na.rm = TRUE)` meters. 


Fixing the temperatures

```{r}
met <- met[temp>-40][order(temp)]
head(met)
```

## Step 5: Checking against external data source.

```{r}
met <- met[temp>-15][order(temp)]
summary(met[, .(lat, lon, wind.sp, temp, elev)])
```

## Step 6: Summarizing everything 

```{r}
met[elev==max(elev,na.rm=TRUE)][,summary(elev)]

met[elev==max(elev,na.rm=TRUE)][,summary(wind.sp)]

met[elev==max(elev,na.rm=TRUE)][,summary(temp)]
```
Computing correlations

```{r}
met[elev == max(elev, na.rm = TRUE), .(
  temp_wind = cor(temp, wind.sp, use = "complete"),
  temp_hour = cor(temp, hour, use = "complete"),
  temp_day = cor(temp, day, use = "complete"),
  wind_day = cor(wind.sp, day, use = "complete"),
  wind_hour = cor(wind.sp, hour, use = "complete")
)]

#This creates a new data table and not modifying the dataset itself

```
## Step 7: Exploratory graphs

```{r}
hist(met$elev, breaks=100)

hist(met$temp)

hist(met$wind.sp)
```

Look at where the weather station with highest elevation is located.

```{r}
elev <- met[elev==max(elev,na.rm=TRUE)]

library(leaflet)

leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```


Look at the time series of temperature and wind speed at this location

```{r}
library(lubridate)

elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)

elev <- elev[order(date)]
head(elev)
```

With the date-time variable we can plot the time series of temperature and wind speed.

```{r}
plot(elev$date, elev$temp, type='l')

plot(elev$date, elev$wind.sp, type='l')
```


There appears to be a daily cyclic pattern in the temperature and windspeed changes at this location. 
