---
title: "Lab 5"
author: "Christina Lin"
date: "9/24/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1: Read the Data

```{r}
library(data.table)
library(tidyverse)

# Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]

```


```{r}
if (!file.exists("met_all.gz")){
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)}

met <- data.table::fread("met_all.gz")

```


Merging the data

```{r}
met <- merge(
  x = met,
  y = stations,
  all.x = TRUE, all.y = FALSE,
  by.x = "USAFID", by.y = "USAF"
)
```


# Question 1: Representative station for the US

Collapse the data set so there is one representative measurement per station using averages

```{r}
station_averages <- met[,.(
  temp = mean(temp, na.rm=TRUE),
  wind.sp = mean(wind.sp, na.rm = TRUE),
  atm.press = mean(atm.press, na.rm = TRUE),
  lat = lat,
  lon = lon
), by = USAFID]
```


Now we need to identify the median per variable

```{r}
medians <- station_averages[,.(
  temp_50 = quantile(temp, probs = .5, na.rm = TRUE),
  wind.sp_50 = quantile(wind.sp, probs = .5, na.rm = TRUE),
  atm.press_50 = quantile(atm.press, probs = .5, na.rm=TRUE),
  lat = lat,
  lon = lon
)]

medians

```

Now we can find the stations that are closest to these
#which.min() returns the index of the observation

```{r}
station_averages[, temp_dist := abs(temp - medians$temp_50)]
median_temp_station <- station_averages[order(temp_dist)][1]
median_temp_station
```


The station that is closest to the median temperature is `r median_temp_station$USAFID`.


```{r}
station_averages[, wind.sp_dist := abs(wind.sp - medians$wind.sp_50)]
median_wind_station <- station_averages[order(wind.sp_dist)][1]
median_wind_station
```

The station that is closest to the median wind speed is `r median_wind_station$USAFID`.




```{r}
station_averages[, atm.press_dist := abs(atm.press - medians$atm.press_50)]
median_atm_station <- station_averages[order(atm.press_dist)][1]
median_atm_station
```

The station that is closest to the median atmospheric pressure is `r median_atm_station$USAFID`.


The representative/median stations for temperature, wind speed, and atmospheric pressure do no coincide.



# Question 2: 

First, need to recover the state variable by merging

```{r}
station_averages <- merge(x = station_averages, y = stations, by.x = "USAFID", by.y = "USAF", all.x = TRUE, all.y = FALSE)

```


```{r}
station_averages[, temp_50 := quantile(temp, probs= .5, na.rm = TRUE), by = STATE]
station_averages[, wind_50 := quantile(wind.sp, probs= .5, na.rm = TRUE), by = STATE]
station_averages[, atm_50 := quantile(atm.press, probs= .5, na.rm = TRUE), by = STATE]
```


```{r}
station_averages[, eudist := sqrt(
  (temp - temp_50)^2 + (wind.sp - wind_50)^2 + (atm.press - atm_50)^2
)]

state_median_station <- station_averages[, .SD[which.min(eudist)], by = STATE]
state_median_station[, .(STATE, USAFID, lat, lon)]

```

This table shows the representative station of each state, calculated by using the lowest euclidean distance for temperature, wind speed, and atmospheric pressure. 




# Question 3

Finding the station at the middle of the state can be done by finding the station with median latitude and longitude in each state. 

```{r}

```



# Question 4

Going back to the met data set
```{r}
met[, state_temp := mean(temp, na.rm=TRUE), by = STATE]
met[, temp_cat := fifelse(state_temp < 20, "low-temp",
                          fifelse(state_temp < 25, "mid-temp","high-temp"))]
```

Let's make sure that we don't have NAs

```{r}
table(met$temp_cat, useNA= "always")
```

Now let's summarize

```{r}
tab <- met[, .(
  N_entries = .N,
  N_NA = sum(is.na(temp)),
  N_stations = length(unique(USAFID)),
  N_states = length(unique(STATE)),
  temp_mean = mean(temp, na.rm = TRUE),
  wind.sp_mean = mean(wind.sp, na.rm = TRUE),
  atm.press_mean = mean(atm.press, na.rm = TRUE)
),by = temp_cat]


#this prints out a pretty table depending on the format of the document (prints a markdown table)
knitr::kable(tab)
```

